Pre- and Post-Debate Democratic Primary Data: Twitter, Google Trends, and Polls

### 2019-12-11 Paper/paper-draft-Padgett-Thiam.Rmd
fix formatting and typos

@@

@@

### Data exploration

Because our project was exploratory in nature, we spent significant time exploring each data source.

Create a function to tabulate the count and proportion of tweets by candidates.

##### Create functions
```{r, eval=FALSE}
type_var  <- unlist(map(pre_debate, class))

@@

A visual of the proportion of tweets pre and post debate

##### Graphing
###### Pre-debate number of tweets

@@

\newpage
Figure 2 shows the post- debate tweets mentions by candidate. As expected, lot more tweets mentions occured after the debate, about 5 times more than pre- debate. Here we noticed that the ranking change with Harris and Yang in the lead along with Sanders and tweets mentions close to 150,000.

@@

Finally, we examined the relationships between the polling data and the Twitter and Google Trends data post-debate, as well as the relationships between pre- and post-debate changes across data sources.

@@

@@

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Twitter and Polling Data"}
print(graphing_data)
pre <-
graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Pre-debate") %>%
  mutate(Pre_twitter=Percentage, Pre_polling=Percentage1, Pre_gtrends=Percentage2)

post <-
  graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate) %>%
  filter(Pre_post=="Post-debate") %>%
  mutate(Post_twitter=Percentage, Post_polling=Percentage1, Post_gtrends=Percentage2)

@@

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Google Trends and Polling Data"}
change_data %>%
  group_by(Pre_post, Candidate) %>%
  mutate(change_twitter = ((Post_twitter - Pre_twitter) / Pre_twitter) * 100, 
         change_polling = ((Post_polling - Pre_polling) / Pre_polling) * 100, 
         change_gtrends = ((Post_gtrends - Pre_gtrends) / Pre_gtrends) * 100) %>%
  ggplot() +
  geom_point(mapping = aes(x=change_gtrends, y=change_polling, color=Candidate), size=4) +
  labs(x="Percent Change in Google Trends Data", 
       y="Percent Change in Polling Data")
```
When looking at the pre- and post-debate data from each source side-by-side, we can see that, while candidates who are polling higher have higher numbers of tweets and Google searches (for the most part), there is no clear relationship between the post-debate changes in polling numbers and the post-debate changes in Twitter and Google Trends data. Regardless of polling numbers, all candidates had higher Twitter positivity ratings and higher numbers of Google searches post-debate than pre-debate.

>>>>

All data used to create the tables, figures, and shiny app in this paper can be found on our Github repository at https://github.com/znpadgett/surv727_padgett_thiam.

@@

### Data Exploration

@@

##### Create functions
Create a function to tabulate the count and proportion of tweets by candidates.

```{r, eval=FALSE}
type_var  <- unlist(map(pre_debate, class))

@@

##### Graphing
A visual of the proportion of tweets pre and post debate

@@

Figure 2 shows the post- debate tweets mentions by candidate. As expected, lot more tweets mentions occured after the debate, about 5 times more than pre- debate. Here we noticed that the ranking change with Harris and Yang in the lead along with Sanders and tweets mentions close to 150,000.

@@

Finally, we examined the relationships between the polling data and the Twitter and Google Trends data post-debate, as well as the relationships between pre- and post-debate changes across data sources. We looked at these results for the 5 most popular candidates.

@@

Figure 5 shows that there is no clear relationship between the post-debate twitter sentiment and the polling data. All candidates have similar sentiment scores, regardless of polling percentage.

@@

In Figure 6, we can see a very slight relationship between the post-debate Google Trends data and polling data, in that candidates polling higher generally had higher Google trends hit rates.

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Twitter and Polling Data"}
pre <-
graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate, Data) %>%
  filter(Pre_post=="Pre-debate") %>%
  mutate(Pre_twitter=Percentage, Pre_polling=Percentage1, 
         Pre_gtrends=Percentage2)

post <-
  graphing_data  %>%
  select(Pre_post, Candidate, Percentage, Percentage1, Percentage2) %>%
  group_by(Pre_post, Candidate, Data) %>%
  filter(Pre_post=="Post-debate") %>%
  mutate(Post_twitter=Percentage, Post_polling=Percentage1, 
         Post_gtrends=Percentage2)

@@

Figure 7 shows the percent change in Twitter sentiment and percent change in vote share between the two time periods (pre- and post-debate). There is no clear relationship between the change in vote share and the change in Twitter sentiment.

```{r, tidy=TRUE, fig.cap="Comparison of Changes in Google Trends and Polling Data"}
change_data %>%
  group_by(Pre_post, Candidate) %>%
  mutate(change_twitter = ((Post_twitter - Pre_twitter) / 
                             Pre_twitter) * 100, 
         change_polling = ((Post_polling - Pre_polling) / 
                             Pre_polling) * 100, 
         change_gtrends = ((Post_gtrends - Pre_gtrends) / 
                             Pre_gtrends) * 100) %>%
  ggplot() +
  geom_point(mapping = aes(x=change_gtrends, y=change_polling, 
                           color=Candidate), size=4) +
  labs(x="Percent Change in Google Trends Data", 
       y="Percent Change in Polling Data")
```

Figure 8 shows the percent change in Google Trends hit rate and percent change in vote share between the two time periods. Again, there is no clear relationship between the change in the two data sources.

In conclusion, when looking at the pre- and post-debate data from each source side-by-side, we can see that, while candidates who are polling higher have higher numbers of tweets and Google searches (for the most part), there is no clear relationship between the post-debate changes in polling numbers and the post-debate changes in Twitter and Google Trends data. Regardless of polling numbers, all candidates had higher Twitter positivity ratings and higher numbers of Google searches post-debate than pre-debate.
